%% ICSE 2027 — Research Track submission (double-blind)
%% Format: IEEE two-column conference. Swap the documentclass if ICSE 2027
%% mandates the ACM acmart format (the section files are class-agnostic).
\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts

% ---------- packages ----------
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{array}
\usepackage{xcolor}
\usepackage{microtype}
\usepackage[hidelinks]{hyperref}
\usepackage{balance}
\usepackage{xspace}

% ---------- convenience macros (names are PLACEHOLDERS; see ICSE/README.md) ----------
\newcommand{\sys}{\textsc{Remedy}\xspace}         % working system name (placeholder)
\newcommand{\hit}[1]{Hit@#1}
\newcommand{\real}{\textsc{JiraInc}\xspace}       % real dataset short tag (placeholder; confirm in Datasets)
\newcommand{\ob}{\textsc{OB}\xspace}              % Online Boutique (synthetic)
\newcommand{\otel}{\textsc{OTel}\xspace}          % OpenTelemetry Demo (synthetic)

\begin{document}

\title{Capability-Gated Agentic Incident Triage:\\
Matching Live Incidents to Resolved Tickets from Telemetry to the Real World}

% --- double-blind: anonymized authors ---
\author{\IEEEauthorblockN{Anonymous Author(s)}
\IEEEauthorblockA{Submitted for double-blind review --- ICSE 2027}}

\maketitle

\begin{abstract}
On-call engineers triage a steady stream of alerts, and their most valuable
signal is institutional memory: how similar incidents were diagnosed and
resolved in the past. Surfacing the right prior incident for a live one is,
however, hard, and most automated-triage research is evaluated only on synthetic
fault-injection telemetry, leaving open how such systems behave on real
engineer-written incidents. We present an agentic incident-triage system that,
given a window of telemetry (or a textual incident report), retrieves the most
similar resolved incidents from a memory of past tickets and decides whether the
window warrants paging an engineer. The system fuses dense, lexical, and
knowledge-graph retrieval; adapts which reasoning steps it runs to the evidence
that is actually available, from rich telemetry to text alone; escalates
expensive steps only when it is uncertain; and suppresses duplicate pages across
the lifetime of one incident. We evaluate on three datasets of increasing
realism. Two we collect and release: microservice-telemetry datasets built by
injecting controlled faults into open-source applications and capturing the
resulting logs, metrics, distributed traces, and cluster events alongside
incident tickets and per-window ground-truth labels---a telemetry-to-ticket
pairing we are not aware of in any public dataset. The third is a large
real-world corpus of resolved Apache Jira incidents (38{,}642 tickets across 24
projects), drawn from the publicly available World of Logs collection. On real
incidents, the fusion retriever places the correct prior incident in its top five
$97\%$ of the time, outperforming dense ($0.91$), lexical ($0.73$), graph
($0.31$), and a retrieval-augmented large-language-model baseline ($0.86$), all
significant under multiple-comparison correction; the end-to-end agent reaches
$0.96$ while its capability gating cuts triage cost by up to $78\%$ at equal
accuracy. A controlled synthetic-versus-real comparison further shows that the
ticket-worthiness classification that dominates synthetic benchmarks is trivially
saturated yet unlearnable on real incidents, relocating the genuine difficulty to
retrieval --- a finding that reframes how operational-triage systems should be
evaluated. Our two telemetry datasets, code, evaluation harness, and all results
are publicly released.
\end{abstract}

\begin{IEEEkeywords}
incident triage, AIOps, information retrieval, LLM agents, empirical software engineering
\end{IEEEkeywords}

\input{sections/01-introduction}
% \input{sections/02-background}   % TODO: not yet written
\input{sections/03-approach}
\input{sections/04-datasets}
\input{sections/05-setup}
\input{sections/06-results}
\input{sections/07-discussion}
\input{sections/08-threats}
% \input{sections/09-related}      % TODO: author-written
\input{sections/10-conclusion}

\balance
\bibliographystyle{IEEEtran}
\nocite{*} % TEMP: include placeholder bib entry so the draft compiles; remove once real \cite{} calls exist
\bibliography{refs}

\end{document}
