% === Background and Motivating Example ===
\section{Background and Motivating Example}
\label{sec:background}

Consider an engineer on call for a microservice application who is paged when
checkout requests begin to fail. The pressing question is not how to repair the
system from first principles but \emph{what is happening} and \emph{whether it
warrants action}. In a fully instrumented deployment the evidence is rich yet
noisy: the cluster reports that a message broker's pod has become unavailable, the
services that consume from it begin logging errors, distributed traces show their
requests failing downstream, and an availability alert fires. Assembling these
fragments into a diagnosis is exactly the slow, expertise-bound step that triage
is meant to accelerate.

The shortcut a seasoned engineer takes is \emph{recognition}. This pattern, where a broker outage stalls its consumers and ultimately manifests as checkout failures, has occurred before, and a previously resolved ticket already records how it was diagnosed and fixed. Retrieving that ticket transforms an open-ended investigation into a near-instant analogy. In this sense, the institutional memory encoded in resolved incidents provides the most direct path from symptom to resolution, provided the relevant past incident can be retrieved.

Two characteristics of operational incident triage fundamentally shape our system design. First, the \emph{same underlying incident can arrive with vastly different evidence}. The broker failure above may be accompanied by logs, metrics, traces, and cluster events; in another team's issue tracker, the identical failure may be reported as nothing more than a sentence---``consumers stalled after the broker restarted, backlog growing''---without any telemetry. Yet the retrieval task is identical in both cases: identify the most relevant previously resolved incident. A system that depends on complete observability therefore fails precisely when historical knowledge is needed most. Second, an incident is \emph{not an instantaneous event}. It evolves over a sequence of observation windows, from early warning signs through active failure to eventual recovery. Treating each window independently produces repeated alerts for what is, in reality, a single evolving incident.

\smallskip
\noindent\textbf{Problem formulation.}
We formulate incident triage as adaptive retrieval over a time-aware incident memory under heterogeneous evidence. Let
$\mathcal{W}=\{w_1,\ldots,w_n\}$ denote a time-ordered stream of observation
windows, where each window
\[
w_i=(t_i,\mathbf{x}_i,e_i)
\]
is observed at time $t_i$, contains a textual summary $\mathbf{x}_i$, and may
include an arbitrary subset of auxiliary evidence
$e_i\subseteq\{\text{metrics},\text{logs},\text{traces},\text{events}\}$.
Consequently, different windows may expose different evidence modalities,
including text alone.

The retrieval memory
$\mathcal{M}=\{m_1,\ldots,m_N\}$ consists of previously resolved incidents.
Each incident
\[
m_j=(\mathbf{d}_j,\mathcal{E}_j,\tau_j)
\]
contains its textual description $\mathbf{d}_j$, a set of extracted entities
$\mathcal{E}_j$ (e.g., services, components, error classes, and symptoms), and
its resolution time $\tau_j$. To prevent temporal leakage, only incidents
resolved before the current observation are searchable; for a window observed at
time $t_i$, the eligible memory is
\[
\mathcal{M}_{<t_i}=\{m_j\in\mathcal{M}\mid \tau_j<t_i\}.
\]

Given a window $w_i$, the objective is to retrieve the top-$k$ most relevant
historical incidents from $\mathcal{M}_{<t_i}$ while simultaneously deciding
whether the window warrants an engineer page. Since consecutive windows often
belong to the same underlying operational event, the system must additionally
associate related windows into a single incident so that only one page is issued.
The challenge is to solve this retrieval and triage problem consistently across
the full spectrum of evidence regimes, from text-only reports to richly
instrumented deployments.
