% Section 3 — Approach: the TCH cascade
% Page budget: ~2 pages in acmart sigconf two-column (~1300-1500 words).
% Source-of-truth: short-technical/sections/{02-dataset-construction, 04-pipelines, 05-cascade}.tex
%
% Editorial discipline (per author direction):
%   - Describe the IDEA, not the implementation primitive (no ScaleDeployment,
%     SetEnv, RestartPods, RecordOnly, NetworkChaos, kubectl, YAML, etc.).
%   - No numbers that are downstream RESULTS (those live in §5): coefficient
%     values, ablation deltas, per-pipeline metric values, latency numbers,
%     standalone precision/recall figures, etc.
%   - No leakage from other sections.

\section{Approach}
\label{sec:approach}

\subsection{Dataset and labels}
\label{sec:approach:dataset}

The workload is Google's Online Boutique~\cite{onlineboutique}, a ten-service e-commerce demo whose services are written in six languages and communicate over remote procedure calls with a cache backing the cart service. We forked the upstream code and added production-realistic telemetry under a single binding rule: \emph{anything a real production service would not emit must not be added}. Concretely, we reject any field that names our scenario taxonomy---no scenario identifiers in span attributes, no fault-type strings in log fields, no alert names tied to our incident families. The discipline makes debugging harder but the resulting telemetry indistinguishable from what an instrumented production fleet would emit, so the evaluation that follows describes the production task rather than a fictional easier surrogate.

Three telemetry channels are added to every service. The logging channel emits a structured record per remote-procedure call, plus a structured error record at every dependency boundary capturing what was being called, with what operation, with what error class, and at what retry attempt. A small set of business events records domain-level signal (cart-size changed, order placed, payment charged) without ever encoding scenario identity. The tracing channel attaches errors to the span where they occurred and adds manual child spans for dependency calls with standard semantic attributes. The metrics channel exports the rate, error, and duration of every handler, together with per-dependency client breakdowns and standard runtime gauges from each language's default exporter.

Faults are injected through four conceptual mechanisms that collectively define a taxonomy of 27 scenario families. \emph{Service removal} takes a chosen service offline; this backs the eight hard-outage families and the cart-dependency outage. \emph{Behavioral perturbation} changes the configuration of a still-running service (slowing it, biasing its responses, raising its traffic rate); this backs latency-injection and traffic-pressure scenarios. \emph{Pod-level disruption} terminates and replaces pods to model graceful redeploys, single-pod restarts within a healthy replica set, and pathological flapping. \emph{Network-level disruption} delays, drops, or partitions traffic between specific service pairs via a chaos-engineering tool. A fifth, trivial mechanism is passive observation under normal traffic, which produces the negative-class baseline windows.

Each fault produces an \emph{incident episode} with three associated \emph{telemetry windows}: a pre-fault baseline captured just before injection, an active-fault window during the injection itself, and a recovery window after the fault is cleared. Long-running scenarios such as slow memory leaks add a fourth, longer observation window. Every window carries five categories of label that downstream training and evaluation consume.

The first is the triage class, one of \emph{ticket-worthy}, \emph{borderline}, or \emph{noise}. We use three classes rather than two because forcing borderline cases into a binary task biases the evaluation; the third class lets us measure how the model handles genuinely ambiguous windows without penalizing it for refusing to commit. The second is, for ticket-worthy windows, the metadata the corresponding Jira ticket would carry: a severity (minor, major, or critical), the list of affected components, and a coarse reason class. The third is the memory-match ground truth---the list of past Jira tickets whose scenario family matches the window's, restricted to tickets time-visible at the window's timestamp and originating from a different fault-injection run so that the model cannot trivially recover the gold by recognizing its own run. A Boolean \emph{novelty} label fires when this list is empty. The fourth is a fixed-dimensional vector of pre-computed numeric features summarizing the window's logs, traces, metrics, and Kubernetes-state counters, including both absolute moving averages and active-versus-pre-fault deltas; these are the inputs the numeric-feature pipelines consume. The fifth is a free-text evidence summary that interleaves a one-line description of metric anomalies, a small set of characteristic log lines extracted by a baseline-versus-fault diff, and the names of any firing alerts; this text stands in for the engineer's view of the window when retrieving past tickets.

A separate 347-ticket humanized Jira memory corpus paired with the telemetry, plus a 110-ticket distractor pool used to test robustness, are described in \S\ref{sec:background:lab}. The 6{,}720 windows are split chronologically and family-stratified into train, validation, and test partitions; every scenario family appears in all three partitions via disjoint runs, modeling a mature team's operational view. A complementary leave-one-family-out cross-validation, in which an entire family is hidden from training and evaluation switched to that family alone, tests the harder orphan-family regime.

\subsection{Seven pipelines and the Pareto frontier}
\label{sec:approach:panel}

We train seven diagnostic pipelines, each representing a different inductive bias on the joint telemetry-Jira problem. Their union is what the cascade has available to compose; the rest of this subsection describes each, organized by what the cascade ends up consuming from it.

\paragraph{Histogram gradient-boosted classifier on numeric features.} The strong triage baseline. A tree ensemble of three hundred depth-eight regression trees is fit to the pre-computed numeric feature vector that summarizes each window's logs, traces, metrics, and Kubernetes-state counters, and emits a calibrated probability that the window is worth a ticket. Class weights are balanced to compensate for the natural skew between ticket-worthy and noise windows. The model is scale-invariant, requires no standardization, fits in seconds, and produces no retrieval output. It is included because triage is a fundamentally tabular task on this dataset, and a small tree ensemble is the right inductive bias for non-linear interactions over per-feature value ranges that span several orders of magnitude.

\paragraph{Fine-tuned dense bi-encoder.} The strongest single retriever for top-one precision. A small pretrained sentence transformer encodes the window's free-text evidence and each memory ticket's text \emph{independently} into a low-dimensional unit vector, after which similarity is a cosine product. The encoder is fine-tuned end-to-end with a contrastive ranking objective in which the (window, gold-ticket) pair must dominate every other pair in the batch; each anchor sees two hard negatives mined by sparse retrieval plus one randomly drawn negative. The random-negative slot forces the model to learn semantic discrimination beyond lexical overlap, where a hard-only diet otherwise lets the encoder collapse onto the sparse signal it was trained against. A tiny logistic regression over similarity-derived features (the top-one cosine, the mean of the top-five cosines, and the count of memory documents above a similarity floor) converts the retrieval distribution into a triage probability.

\paragraph{Two hybrid sparse-plus-dense fusion retrievers.} A learned-sparse retriever, a separately-trained dense bi-encoder, and a structural graph score over the same memory corpus are run independently and combined into a single ranked list via rank fusion. The sparse retriever encodes each text into a sparse vocabulary-sized vector whose non-zero coordinates are learned term weights rather than raw frequencies, so that semantically related terms inflate alongside the literal terms a text contains; this catches verbatim matches that the dense encoder misses. The graph retriever scores each memory ticket by structured entity overlap against a knowledge graph populated from the language-model extraction below. The two hybrid variants differ only in how the \emph{query-side} entities are extracted at retrieval time: one variant uses hand-written rules over a canonical vocabulary, the other invokes a language model under a strict JSON schema. The two variants exhibit a striking trade-off that drives the cascade's L2 design (\S\ref{sec:approach:cascade}): the language-model extractor is more precise per entity but produces sparser overlap sets, and inside rank fusion that sparsity manifests as anti-consensus.

\paragraph{Log-sequence transformer trained from scratch.} A two-stage neural model that turns a window's \emph{sequence of log lines} into a single vector. Each line is independently embedded by a frozen pretrained sentence transformer; the resulting sequence of line embeddings is then processed by a small two-layer transformer encoder with a learned classification token, trained from scratch on (window-log-sequence, gold-ticket-text) pairs with the same contrastive objective as the bi-encoder. The frozen line encoder transfers general-text semantics for free, leaving only the sequence-aggregation step for the small training set to learn. This pipeline's standalone retrieval is weak on average but \emph{exceptionally strong} on scenario families whose log signature is stereotyped---in particular product-catalog outages, single-pod restarts in healthy replica sets, and slow-recovery windows---because those windows are distinguished by an ordering pattern that bag-of-words representations cannot see.

\paragraph{Knowledge-graph retriever.} A structural retriever that scores past tickets by weighted entity overlap against a small knowledge graph built once, offline, from the language-model extraction. The graph carries one node per memory ticket plus typed nodes for affected services, components, observable symptoms, error classes, root causes, and resolutions, with edges linking each ticket to the entities it mentions. At retrieval time, the query window's entities are extracted by the same canonical-vocabulary rule extractor used by one of the hybrid variants, and a single parameterized graph query scores each ticket by an additive overlap function that weights error-class matches more highly than service matches, service matches more highly than component matches, and component matches more highly than symptom matches. The retrieval is intentionally broad: a typical query window shares at least one entity with dozens of past tickets, so the top-one slot is genuinely contested, but the right ticket almost always appears somewhere in the top-twenty. The pipeline is in the panel both for its structural signal and for the per-match natural-language citation it can produce alongside each retrieved candidate.

\paragraph{Three-stage language-model agent.} The only pipeline that runs a language model at inference time, and the only pipeline that emits a Boolean novelty flag. Each window is processed by a deterministic three-stage chain. The model first reads the window's evidence text and emits a one-sentence root-cause hypothesis under a strict JSON schema, with the model's chain-of-thought reasoning disabled (the hypothesis stage is treated as extractive and need not deliberate). The cascade then hands the model a small set of cached candidate past tickets from a cheaper retriever, and the model performs an explicit verification stage---chain-of-thought enabled this time---in which it scores each candidate's consistency with the hypothesis on a calibrated zero-to-one scale and marks each candidate as a worthwhile lead or not. When no candidate clears a consistency threshold the model emits a top-level novelty flag. The verification stage is the agent's load-bearing step: it is the only place in the panel that an explicit comparison between hypothesized cause and candidate-ticket cause is performed, and it is where the project's novelty signal originates.

The pipelines are deliberately heterogeneous because no single one dominates the others. The numeric-feature classifier owns the triage axis, the dense bi-encoder wins top-one retrieval precision, the rule-graph hybrid fusion wins top-five retrieval coverage, the log-sequence transformer wins on specific scenario families, the structural retriever produces interpretable citations, and only the language-model agent emits novelty signal at all. The natural alternative---training a single larger model that learns the union of these signals---is unattractive: the cheapest pipeline already saturates the triage axis, the input modalities are deeply heterogeneous (a fixed-dimensional numeric vector, free-text evidence, a variable-length log token sequence, structured graph queries), and the agent's novelty judgment is qualitatively unlike any similarity score so cannot be replicated by tighter encoder training. We therefore compose the seven into a cascade.

\subsection{Cascade architecture}
\label{sec:approach:cascade}

The Tiered Cascade Hybrid (\tch{}) is a four-layer composition organized by per-call cost: a fast triage gate (Layer~1) runs on every window, a retrieval fusion (Layer~2) consumes cached upstream predictions, a novelty check (Layer~3) consumes one cached agent output per window, and a final composition layer (Layer~4) assembles the result. The layers are deliberately independent: the novelty check never reorders the retrieval ranking, the retrieval layer never feeds back into triage, and the triage decision never gates retrieval. This independence is what makes per-layer ablation clean and what allows operators to swap any single pipeline without disturbing the others.

\paragraph{Layer 1: triage gate.} Layer~1 outputs a calibrated probability that the window is worth a ticket. It is a class-balanced logistic regression \emph{stacker} over the per-pipeline triage scores:
\begin{equation}
  P_{L1} \;=\; \sigma\!\left(\sum_{i} w_i\,x_i + b\right),
  \label{eq:l1}
\end{equation}
where $x_i$ is the triage score emitted by the $i$-th pipeline. The stacker is fit via out-of-fold cross-validation so that every training window is scored by a model that never saw it during fit, and class weights are balanced to compensate for the natural skew between ticket-worthy and noise windows. The numeric-feature classifier carries by far the largest learned weight; the remaining pipelines act as small corrections that pay off on borderline-class windows, where the numeric classifier alone is uncertain.

\paragraph{Layer 2: retrieval fusion.} Layer~2 produces a ranked list of five past tickets. It uses two different combination rules because the top-1 slot and the remaining four slots optimize different objectives: the top-1 slot rewards a confident consensus pick, whereas positions 2--5 reward diverse coverage. For the top-1 slot, we take the dense retriever's three highest-ranked candidates as an \emph{anchor pool} and re-rank them by how strongly the other retrievers concur,
\begin{equation}
  \mathrm{overlap}(c) \;=\; \sum_{r \in \mathcal{V}} \mathbf{1}[c \in \mathrm{top3}(r)]\,\bigl(3 - \mathrm{rank}_r(c)\bigr),
  \label{eq:l2-overlap}
\end{equation}
where the voter set $\mathcal{V}$ includes the two hybrid-fusion retrievers and the log-sequence transformer. A candidate that several independent retrievers agree on is more trustworthy than the dense retriever's own first guess; in the absence of any overlap, the dense retriever's top candidate is used unchanged.

For positions 2--5 we apply Reciprocal Rank Fusion~\cite{cormack2009rrf}:
\begin{equation}
  \mathrm{RRF}(c) \;=\; \sum_{r \in \mathcal{R}} \frac{1}{60 + \mathrm{rank}_r(c)},
  \label{eq:l2-rrf}
\end{equation}
summed over a retriever set $\mathcal{R}$ comprising the dense retriever, one of the two hybrid-fusion variants, the log-sequence transformer, and the knowledge-graph retriever. Candidates are sorted by their RRF score; the anchor (already at position~1) is removed; the next four candidates fill positions 2--5. We chose Reciprocal Rank Fusion deliberately over learned or score-weighted alternatives: it requires no per-retriever score calibration and is invariant to monotone score transformations, which makes it robust to retrievers whose confidence distributions differ.


\paragraph{Layer 3: three-signal novelty check.} Layer~3 produces the Boolean novelty flag that distinguishes recurring incidents from genuinely new ones. It is a disjunction of three independent signals,
\begin{equation}
\small
\mathtt{is\_novel}
=
\mathtt{agent\_novel}
\lor
(\max_i \mathrm{conf}_i < 0.5)
\lor
(P_{\text{learned}}(\mathrm{novel}) \geq 0.5),
\label{eq:l3}
\end{equation}
\normalsize
each of which captures a different reason the cascade might be wrong about a window having a past match. The first disjunct is the agent's own verify-stage judgment, expensive but specific. The second is a threshold over the maximum per-retriever confidence: if no retriever is even half-confident in any candidate, the cascade has effectively no past-match hypothesis, regardless of what the agent says. The third is a learned classifier over coarse window metadata---window type, scenario family, service name, a flag for intentionally-difficult cases, the maximum retrieval confidence, and the triage probability---trained again via out-of-fold cross-validation; it captures regularities the other two miss, such as that pre-fault baseline windows are novel by construction and recovery windows are not. We use disjunction rather than a weighted combination because the three signals capture qualitatively different forms of evidence and operators benefit from being able to tune a single learned-signal threshold without redesigning the others.

\paragraph{Layer 4: composition.} Layer~4 packs the calibrated triage probability, the ranked top-five past tickets, and the novelty flag into the final output and derives a coarse ticket-worthy-or-noise decision from the triage probability via the Layer-1 threshold. No further computation occurs.

The four-layer composition is what makes seven Pareto-incomparable pipelines compose into a system that beats every individual baseline on every metric. The layers do not interact: each is replaceable in isolation, each is ablatable in isolation, and the cascade's behavior on any given window can be traced layer by layer back to the upstream pipeline outputs it consumes.
