% === Experimental Setup ===
\section{Experimental Setup}
\label{sec:setup}

\subsection{Research Questions}
\label{sec:setup-rqs}

Our evaluation is organized around four questions:

\begin{itemize}
  \item[\textbf{RQ1}] \textbf{(Retrieval).} How accurately can a live incident be
  matched to the most similar resolved incident in memory, and does fusing dense,
  lexical, and graph retrieval outperform each in isolation, as well as strong
  prior-art and LLM-based retrievers?
  \item[\textbf{RQ2}] \textbf{(End-to-end triage).} Does the capability-gated
  system retrieve effectively end to end, and does the \emph{same} system hold up
  as the available evidence degrades from full telemetry to text alone?
  \item[\textbf{RQ3}] \textbf{(Cost of adaptive execution).} What does the system's
  capability gating buy? Specifically, how much triage cost does it save relative
  to running every step on every window, at equal accuracy, and which components
  carry the result?
  \item[\textbf{RQ4}] \textbf{(Synthetic vs.\ real).} How do triage and retrieval
  behave on synthetic versus real incidents, and what does the contrast reveal
  about where the difficulty truly lies?
\end{itemize}

\subsection{Systems and Baselines}
\label{sec:setup-baselines}

We compare our hybrid retriever and full system against three groups of baselines,
all evaluated under an identical memory, time-aware visibility, and gold contract.

\noindent\emph{Constituent retrievers.} To show that fusion is more than the sum
of its parts, we evaluate each of our retrievers alone: the fine-tuned dense
bi-encoder, the learned sparse retriever, and the knowledge-graph retriever.

\noindent\emph{Prior-art retrieval.} We add widely used, off-the-shelf retrievers
applied zero-shot: three strong dense embedding models (BGE, E5, and
MPNet)~\cite{xiao2024cpack,wang2022e5,song2020mpnet}, classical lexical search
(TF--IDF and BM25~\cite{robertson2009bm25}), and a cross-encoder
re-ranker~\cite{nogueira2019passage} applied over the dense candidates.

\noindent\emph{Retrieval-augmented LLM.} As a contemporary point of comparison,
an instruction-tuned language model (Qwen2.5-7B~\cite{qwen2025}) re-ranks the
dense top candidates for each query; following common practice for LLM baselines,
we run it on a fixed 500-incident sample per dataset.

\noindent\emph{Triage classifiers.} For the triage decision (RQ4) we additionally
compare standard classifiers over the numeric telemetry features---gradient-boosted
trees, random forests, and logistic regression---wherever such features exist.

\subsection{Metrics}
\label{sec:setup-metrics}

We measure two capabilities separately. \emph{Retrieval} is scored with
Hit@$k$---the correct prior incident appears in the top $k$, for
$k\in\{1,5,10\}$---and mean reciprocal rank (MRR), averaged over queries that have
at least one ground-truth match; on the real dataset we report both the coarse
and strict matching relations of \S\ref{sec:datasets-wol}. \emph{Triage} is a
calibrated classification problem, so beyond accuracy we report the area under the
precision--recall curve, the area under the ROC curve, expected calibration
error, and precision at a fixed $5\%$ false-positive rate---the operating point an
alerting system actually runs at. We further report novelty detection (precision
and recall of flagging incidents with no precedent) and, operationally, pages per
incident, which captures whether duplicate alerts are suppressed. Finally,
\emph{cost} is measured per window as wall-clock time and language-model token
usage, relative to a counterfactual cascade that runs every step on every window.

\subsection{Protocol and Implementation}
\label{sec:setup-protocol}

\noindent\emph{Splits.} The synthetic datasets are split by holding out whole
fault families, and the real dataset by holding out two entire project domains
(\S\ref{sec:datasets}); all numbers are reported on the held-out test split.

\noindent\emph{Statistics.} Every reported difference uses a paired bootstrap
($1{,}000$ resamples, $95\%$ confidence intervals) in which the same resampled
windows are scored for both systems, preserving per-window correlation; a
difference is significant when its interval excludes zero. Because we run many
comparisons, we control the false-discovery rate across the whole family of tests
with the Benjamini--Hochberg procedure and report corrected $q$-values.

\noindent\emph{Gold validation.} Because the real dataset's matches are inferred
rather than human-labelled, we additionally validate them with an LLM-as-judge: a
language model rates the relevance of each gold match and of a random control, and
we report the gap between the two as evidence that the labels carry signal.

\noindent\emph{Implementation.} The dense retriever is a compact MiniLM
bi-encoder fine-tuned for five epochs with hard negatives mined by lexical search;
the hybrid fuses each retriever's top candidates by reciprocal rank fusion. Graph
entities are extracted from tickets by a small language model and matched by
entity overlap. The retrieval-augmented and judge models run offline with
deterministic (greedy) decoding. Every experiment uses a fixed random seed, and
full configurations and seeds are released with the artifact.

\noindent\emph{Compute environment.} All experiments were executed on the
Trillium cluster, a system administered by the Digital Research Alliance of
Canada.\footnote{\url{https://alliancecan.ca}} Each training and sampling job was
allocated a single NVIDIA H100 GPU (80\,GB), 8 CPU cores, and 32\,GB of host
memory.
