% Section 6 — Cross-app generalization
% Page budget: ~1 page (the differentiating contribution)
%
% This section is the strongest novelty axis of the paper. Many AIOps papers
% report numbers on one app; almost none demonstrate transfer to a structurally
% different application. The cross-app result + graded difficulty (L1-L4) +
% multi-gold metrics are the new claim space.
%
% Subsections:
%   6.1 Setup
%       - OpenTelemetry Demo (18 services, Kafka async, 12 languages)
%       - Same harness, same locked cascade, same evaluation protocol
%       - 4 new architectural axes vs OB: Kafka async bridge, 2x trace depth,
%         polyglot span emitters, LLM service
%
%   6.2 Headline cross-app table (L1 single-fault only, for OB comparison)
%       - Two columns: zero-shot (locked cascade) + L1-retrained (refit stacker only)
%       - Same metrics as §5 headline
%
%   6.3 Graded difficulty: L1 → L2 → L3 → L4
%       - New table, no OB analogue
%       - Hit@5 / AllGold@K / PrimaryGold@K / Recall@K per difficulty
%       - Monotone degradation expected; report exact numbers
%
%   6.4 Kafka families (architectural-distance evidence)
%       - Sub-table: per-Kafka-family Hit@5
%       - Discussion: silence-of-downstream-consumer as a novel signal
%
%   6.5 Feature-population gap (honest finding)
%       - 12/94 features populate on OTel Demo (generic telemetry: trace counts, CPU%, memory%)
%       - 82/94 require OB-specific Prometheus metric names + Loki label structure
%       - The cascade structurally generalizes; feature values degrade by source-app instrumentation

\section{Cross-App Generalization}
\label{sec:cross-app}

\todo{The cross-app section. Write after GCP collection completes
      (currently in progress). Until then, this is a structural placeholder.
      The architectural-distance evidence is the most novel piece:
      OTel Demo has Kafka async (no OB analogue), 2$\times$ trace depth,
      polyglot span emitters, and an LLM service. Numbers will populate
      from data/derived/global/2026-06-XX-otel-demo-v1-global/.}

\subsection{Experimental setup}
\label{sec:cross-app:setup}

\todo{Two paragraphs: (a) the OTel Demo (18 services, Kafka async, 12+
      languages, LLM service); (b) what's locked and what changes
      (zero-shot transfer: nothing retrained; L1-retrained: only the
      L1 stacker refit).}

\subsection{Headline transfer results}
\label{sec:cross-app:headline}

\todo{Single table with three columns: OB locked, OTel zero-shot, OTel
      L1-retrained, plus a delta column. Same metric rows as §5
      headline. Numbers TBD pending GCP collection results.}

\subsection{Graded difficulty: L1-L4 multi-fault}
\label{sec:cross-app:graded}

\todo{New table, has no OB analogue. Rows: Hit@5, AllGold@K,
      PrimaryGold@5 (L3 only), Recall@K (averaged over golds).
      Columns: L1 single, L2 concurrent, L3 cascade, L4 compound.
      The expected story: monotone degradation L1 -> L2 -> L3, with
      PrimaryGold@5 on L3 being the most demanding metric.}

\subsection{Kafka families: architectural-distance evidence}
\label{sec:cross-app:kafka}

\todo{Per-Kafka-family Hit@5 sub-table (5 families: broker-outage,
      consumer-lag, consumer-crash, partition-rebalance, dead-letter-spike).
      The interpretation: the "silence of downstream consumers" pattern
      is a novel signal not present in any OB scenario. Whether the
      cascade catches it without retraining is the key claim.}

\subsection{Feature-population gap (honest finding)}
\label{sec:cross-app:feature-gap}

\todo{The 12/94 vs 82/94 finding. Frame as: "Structural compatibility
      of the schema is preserved (94/94 columns produced); value
      population varies by source-app instrumentation (12/94 from
      generic telemetry, 82/94 require app-specific Prometheus metric
      naming)." This is a publishable cross-app generalization
      finding, not a bug. Belongs to §6 not §8 because it's evidence
      of how the cascade transfers, not a threat to validity.}
