% Section 8 — Threats to validity
% Page budget: ~0.5 pages
%
% Standard ICSE structure: internal / external / construct.
%
% Internal validity:
%   - The bias-avoidance discipline; pre-registered scenario taxonomy
%   - LLM seed fixed; bootstrap CI with seed=42
%   - The 5-fold CV on the L1 stacker prevents window-level leakage
%
% External validity:
%   - Synthetic Jira corpus (mitigated by TAWOS-anchored humanizer redesign)
%   - 100% trace sampling vs production typical 1-10%
%   - Single-cluster single-region
%   - One synthetic app + one cross-app — not five apps in different domains
%
% Construct validity:
%   - Per-window-level labels are scenario-authored, not engineer-adjudicated
%   - Hit@K assumes engineers review serially; real-world consumption may differ

\section{Threats to Validity}
\label{sec:threats}

\todo{Half a page; standard ICSE structure (internal / external / construct).
      Keep it honest: the 100\% trace sampling, synthetic Jira, single-region
      cluster, and cross-app scope (one cross-app, not five) are real
      limitations. The bias-avoidance discipline is the strongest internal-
      validity argument. Cross-reference §6.5 feature-population gap rather
      than re-listing it here.}

\subsection{Internal validity}
\label{sec:threats:internal}

\todo{One paragraph. Bias-avoidance discipline; pre-registered scenario
      taxonomy; fixed bootstrap seed; 5-fold CV on the L1 stacker
      prevents window-level leakage; LLM determinism via temperature=0
      + fixed seed.}

\subsection{External validity}
\label{sec:threats:external}

\todo{One paragraph. List the production-fidelity divergences (100\% trace
      sampling, single cluster, synthetic Jira, single fault per OB run).
      Anchor the synthetic-Jira concern in the TAWOS-anchored humanizer
      redesign. Acknowledge cross-app evidence is one app, not five.}

\subsection{Construct validity}
\label{sec:threats:construct}

\todo{One paragraph. Hit@K assumes serial candidate review; per-window
      labels are scenario-authored not engineer-adjudicated; the time-to-
      diagnose simulation is a back-of-envelope translation.}
