% Section 9 — Related work
% Page budget: ~0.75 pages
%
% Five categories — one paragraph each:
%
%   9.1 Log anomaly detection
%       - Drain, DeepLog, LogAnomaly, MetaLog, etc.
%       - What they target (detection) vs what we target (diagnosis-by-citation)
%
%   9.2 AIOps & incident triage
%       - Microsoft / Google / FB papers on incident clustering and routing
%       - Distinction: those operate on tickets alone; we operate on telemetry+tickets joint
%
%   9.3 Retrieval-augmented generation (RAG)
%       - Lewis et al. RAG; Atlas; ColBERT
%       - Our cascade uses RAG-adjacent techniques but at runtime is not LLM-driven
%
%   9.4 Knowledge graph for software engineering
%       - KG construction from text; LLM-extracted KGs for incident management
%       - Our LLM-extracted Neo4j graph is one of seven pipelines, not the focus
%
%   9.5 Microservice benchmarks
%       - Online Boutique, Train-Ticket, Sock Shop, DeathStarBench
%       - Discuss why OTel Demo is the right cross-app testbed (Kafka async + polyglot)

\section{Related Work}
\label{sec:related}

\todo{Five short paragraphs, one per category. Cite generously — references
      are unlimited at ICSE. Each paragraph ends with one sentence
      distinguishing our work from the cited line.}

\subsection{Log anomaly detection}
\label{sec:related:logs}

\todo{One paragraph. Distinguish detection (their target) from diagnosis-by-
      citation (ours).}

\subsection{AIOps and incident triage}
\label{sec:related:aiops}

\todo{One paragraph. Cite the major industrial incident-clustering and
      routing papers. Distinguish their ticket-only operation from our
      telemetry+ticket joint approach.}

\subsection{Retrieval-augmented generation and dense retrieval}
\label{sec:related:rag}

\todo{One paragraph. RAG, Atlas, ColBERT, sentence-transformers, MS-MARCO
      cross-encoder reranker. Distinguish RAG (generation-time LLM) from
      our cascade (training/preprocessing-time LLM only; runtime is
      classical).}

\subsection{Knowledge graphs from operational text}
\label{sec:related:kg}

\todo{One paragraph. KG construction from text + LLM-extracted graphs.
      Distinguish our use of the KG as one of seven pipelines from work
      where the KG is the main contribution.}

\subsection{Microservice benchmarks}
\label{sec:related:benchmarks}

\todo{One paragraph. Online Boutique, Train-Ticket, Sock Shop,
      DeathStarBench. Justify the OTel Demo choice for cross-app:
      Kafka async + polyglot + LLM service give four orthogonal
      architectural-distance axes.}
