% === Conclusion ===
\section{Conclusion}
\label{sec:conclusion}

Incident triage is, at its core, an exercise in analogy: the fastest route from a
new alert to a resolution is recognizing which past incident it resembles. We
built \sys, a capability-aware system that operationalizes this view, retrieving
similar resolved incidents by fusing dense, sparse, and graph signals, adapting
which reasoning steps it runs to the evidence each window carries, and suppressing
duplicate pages across an incident's lifetime. Across a realism spectrum from fully
instrumented synthetic services to real Apache Jira incidents, hybrid retrieval
locates the right prior incident in its top five $97\%$ of the time on real data,
beating strong dense, lexical, graph, and LLM baselines, while capability gating
matches that quality at up to $78\%$ lower cost and degrades gracefully to
text-only operation. Our broadest finding is a caution about evaluation: the
ticket-worthiness classification that dominates synthetic benchmarks is trivially
solved on injected faults yet unlearnable on real ones, so it measures the data
rather than the method, while the difficulty that transfers lies in retrieval. We
hope the datasets, harness, and baselines we release help the community study
triage where it is genuinely hard.

\section*{Data Availability}

All source code, the evaluation harness, the collected telemetry datasets, and
the complete result set, including the prior-art and LLM baselines, ablations, and
significance tests, are released to support reproduction. The datasets are openly
available at Zenodo, \url{https://zenodo.org/records/21082064}
(DOI~10.5281/zenodo.21082064), and the source code and evaluation harness at
\url{https://zenodo.org/records/21083181} (DOI~10.5281/zenodo.21083181). Only the derived Apache-project Jira
subset of the World of Logs corpus is redistributed there, under CC~BY~4.0; the full
corpus is publicly available from its original authors under its own terms.
