% Abstract — four paragraphs, ~310 words. Standard ICSE size.
% Industry-pain opening, institutional-memory framing, headline numbers
% in human-readable form, two reviewer-defence negative findings.
% Every number traced to results/{ob,otel-demo,wol}/.../SUMMARY.md.

On-call engineers spend most of their incident-response time not detecting anomalies---automated monitoring handles that---but identifying which past Jira ticket describes the current failure. This lookup is performed by manual keyword search that takes ten to thirty minutes per page, even though the fix almost always already lives in the team's institutional memory. At enterprise scale, the gap compounds into thousands of engineer-hours per year of duplicated diagnostic effort.

We treat the corporate Jira archive as a searchable institutional memory and present a capability-adaptive incident-triage agent that learns from past resolutions to surface the right ticket in milliseconds. For each monitoring window the agent inspects the available evidence---numeric telemetry, structured logs, distributed traces, Kubernetes events, or text alone---and assembles a per-window plan that invokes only the diagnostic methods the evidence warrants, escalating to an expensive language-model verifier only when the cheap path is uncertain. Every decision is replayable from a per-window execution trace.

We evaluate across three datasets spanning the synthetic-to-real spectrum: a synthetic microservices benchmark ($1{,}008$ test windows), the OpenTelemetry Demo ($247$), and $2{,}000$ real Apache Jira tickets across seven projects including Kafka and MariaDB ($304$ windows). The agent matches the always-on cascade's retrieval quality while cutting wall time, language-model tokens, and dollar-equivalent cost by $64.1\%$. On real production tickets, it surfaces the correct past incident as the top suggestion $78.2\%$ of the time and within the top five $83.6\%$---the highest accuracy we observe across any dataset, and, to our knowledge, the first agentic-retrieval result reported on a real Apache Jira corpus. Page suppression converges to one paged ticket per incident on all three datasets.

Two negative findings the framework's instrumentation makes detectable complete the paper: the active evidence-gathering loop's Hit@1 contribution is not statistically distinguishable from zero (paired bootstrap $p = 0.18$), and one of its four tools is significantly harmful on synthetic data ($p = 0.002$). Cross-dataset replication then localises the agent's largest blind spot---a $40\%$ false-novelty rate on synthetic data---to a single overly-eager signal in the language-model verifier: when that signal is structurally absent, the rate collapses by sixty-fold to $0.7\%$.
