% Section 1 — Introduction
% Page budget: ~1 page in acmart sigconf two-column (~750 words).
% Six paragraphs: hook → gap → approach → evaluation regime →
% contributions → roadmap. Every claim cross-checked against
% DOCS/docs8/PAPER-FINDINGS.md + DOCS/docs8/RQ-CLOSURE-TABLE.md +
% per-section SUMMARY.md files. Honest negatives live in §7 and §8
% per the abstract's positive-only framing.

\section{Introduction}
\label{sec:intro}

A modern microservice cluster emits structured logs, distributed traces, numeric metrics, and Kubernetes events at the rate of tens of thousands of records per minute. Most of this stream is uneventful; a small fraction matters; a smaller fraction yet becomes a Jira ticket in the team's institutional memory of past incidents. For the on-call engineer, detecting that something is wrong is largely a solved problem---histogram baselines and learned classifiers over numeric features perform near-perfectly on \emph{``is this anomalous?''} The unsolved problem is what comes next: deciding which past Jira ticket describes the current failure. Today this lookup is performed by manual keyword search, takes ten to thirty minutes per page, and depends on the engineer recalling the right vocabulary. At enterprise scale the gap compounds into thousands of duplicated engineer-hours per year.

Automating the lookup is harder than it appears, for two reasons. First, no public dataset jointly contains rich telemetry and linked Jira issues: the largest log corpora (HDFS, BGL, Thunderbird) have no ticket-side labels; the largest open Jira archive (TAWOS, $458{,}000$ real issues) has no telemetry; and production data from cooperating organizations is generally not shareable for legal and competitive reasons. Second, even when paired data is available, today's diagnostic systems sit at two design extremes. One runs every retriever and every language-model verifier on every window---accurate but expensive in wall time, LLM tokens, and dollars. The other fixes a single pipeline at deployment time and ignores the variety of evidence different incidents present: a CPU saturation looks different in metrics, a cascading Kafka outage looks different in traces, a third-party authentication blip looks different in logs. The choice between cost and coverage is made once at deployment, not per window.

This paper presents a \emph{capability-adaptive incident-triage agent} that makes this choice once \emph{per window}. The agent inspects the available evidence for each window---which modalities are present, at what richness---and emits a per-window plan that invokes only the diagnostic methods the evidence warrants. The cheap retrieval path runs first; expensive retrievers and a language-model verifier escalate only when the cheap path is uncertain. An active evidence-gathering loop can fetch four additional signals on demand (Kubernetes events, expanded trace windows, Prometheus snapshots, peer-incident tickets) when retrieval consensus is low. Every diagnostic invocation is cached; every decision is replayable from a per-window execution trace that records the plan, the gates that fired, the skill outputs, and the budget spent.

We evaluate the agent on three datasets that stress different parts of its design. The first is a controlled synthetic microservices benchmark built from Google's Online Boutique demo: $1{,}008$ test windows across twenty-seven incident scenario families paired with a $347$-ticket humanized Jira memory corpus. The second is the OpenTelemetry Demo, a structurally different polyglot application with $247$ test windows; we use it to probe \emph{zero-shot} transfer, running the same agent against OTel Demo unmodified from its Online-Boutique-trained configuration. The third is World of Logs, a corpus of $9{,}341$ real Apache Jira tickets across seven projects (Spark, Cassandra, HBase, Flink, Ambari, Kafka, MariaDB-Server) with $1{,}648$ test windows; we use it under a \emph{family-held-out} leave-one-domain-out split where the BiEncoder retriever is refit on four families (Spark, Cassandra, HBase, Flink) and the held-out test families (Kafka, MariaDB-Server) are never seen during training. Every headline number carries a $1{,}000$-resample paired bootstrap confidence interval at seed $42$.

\paragraph{Contributions.}
\begin{itemize}
  \item \textbf{Two synthetic incident-triage benchmarks pairing the full telemetry stack with Jira tickets} (\S\ref{sec:evaluation}). Online Boutique and OpenTelemetry Demo each pair Loki logs, Tempo traces, Prometheus metrics, and Kubernetes pod events with humanized Jira memory corpora in a shared timeline schema---to our knowledge the first public benchmarks to do so.
  \item \textbf{A capability-adaptive incident-triage agent} (\S\ref{sec:approach}). The agent emits a per-window diagnostic plan over fifteen registered diagnostic skills---retrieval primitives, cascade-composition steps, and on-demand evidence-fetching tools---gated by the evidence each window presents. Adding a new diagnostic method or evidence modality is a registry edit, not an architectural change.
  \item \textbf{Cost savings without retrieval-quality loss} (\S\ref{sec:results}). On synthetic microservices telemetry the agent matches the always-on cascade's Hit@5 of $0.758$ while spending $64.1\%$ less in dollar-equivalent LLM cost; the savings are robust to a tenfold sweep of the cheap-path confidence threshold and converge to one paged ticket per incident across every dataset.
  \item \textbf{Cross-dataset transfer in two regimes} (\S\ref{sec:cross-app}). The same agent runs zero-shot from its Online-Boutique-trained configuration on the OpenTelemetry Demo and reaches Hit@5 $= 0.756$. Under a family-held-out split on real Apache Jira data, the agent surfaces the correct past incident as its top suggestion $95.8\%$ of the time and reaches Hit@5 $= 1.000$ on the retrieval-evaluable subset---to our knowledge the first agentic-retrieval result reported on a real Apache Jira corpus. Across the three datasets, the agent's Hit@5 dominates a BM25-only baseline by $1.6\times$ on real Apache Jira ($245\times$ at Hit@1) and $8.7\times$ on synthetic microservices.
  \item \textbf{A reproducible measurement harness} (\S\ref{sec:evaluation}). Per-window execution traces, per-skill cost tables, paired-bootstrap delta confidence intervals, and a five-mode tool-use failure-mode taxonomy are emitted for every measurement. Replication is a single command per dataset.
\end{itemize}

The remainder of the paper is organized as follows. Section~\ref{sec:background} fixes terminology and surveys the relevant literature. Section~\ref{sec:approach} describes the agent. Section~\ref{sec:evaluation} defines the evaluation methodology---datasets, splits, baselines, and statistical envelope. Section~\ref{sec:results} reports within-dataset results; Section~\ref{sec:cross-app} reports cross-dataset transfer. Sections~\ref{sec:discussion}--\ref{sec:related} discuss findings, threats to validity, and related work, before concluding in Section~\ref{sec:conclusion}.
