% Section 6 — Cross-dataset transfer
% Page budget: ~2.5 pages in acmart sigconf two-column (~2200 words).
%
% Source-of-truth verification (every number, no fake info):
%   OTel headline: DOCS/docs8/PAPER-FINDINGS.md L150-162 + results/otel-demo/.
%     Hit@1=0.6471 [0.5641,0.7411], Hit@5=0.7563 [0.6726,0.8333],
%     MRR=0.6880 [0.6132,0.7715], Triage=0.8421 [0.7976,0.8866],
%     pages_per_incident=1.000, suppressions=1, plan_ids=4 of 6,
%     cost savings=65.3% USD.
%   WoL v2 headline: PAPER-FINDINGS L164-205 + results/wol-v2/.
%     Hit@1=0.958, Hit@5=1.000, MRR=0.976 (n_eval=48 agent / 276 cascade),
%     Triage=0.164 (below 0.496 majority-class baseline; dense_only
%     recovers to 0.921), pages_per_incident=1.000, suppressions=510
%     (over 901 multi-ticket cases), plan_ids=2 of 6, BM25 lift 1.64x
%     Hit@5 / 245x Hit@1.
%   Cross-dataset false-novel: PAPER-FINDINGS L399-409: OB 40.6%,
%     OTel 39.7%, WoL v2 0.0%.
%   Transfer regimes: OTel zero-shot from OB (no refit); WoL family-held-
%     out leave-one-domain-out (BiEncoder refit on Spark/Cassandra/HBase/
%     Flink train; Kafka/MariaDB-Server held-out test).
%
% Editorial discipline:
%   - Honest cross-dataset reporting includes the triage-accuracy
%     collapse on WoL v2; it is not the abstract's job to surface this
%     but §6's job is exactly to report what generalises and what does
%     not.
%   - The §5 OB-deep story is referenced rather than re-stated.

\section{Cross-Dataset Transfer}
\label{sec:cross-app}

The agent's design (\S\ref{sec:approach}) was developed against \dsob{}. The within-dataset results (\S\ref{sec:results}) show that the design's structural claims---cost savings without retrieval-quality loss, capability-driven gating, statistically significant load-bearing primitives---hold on the synthetic microservices benchmark. This section asks whether they survive a change of dataset, and reports what generalises, what does not, and what the cross-dataset comparison itself reveals about the agent's behaviour.

We test the agent under two distinct transfer regimes. On the OpenTelemetry Demo, the agent runs \emph{zero-shot}: the dense bi-encoder is not refit, the triage stacker is not refit, the knowledge graph is not rebuilt, and only the per-dataset profile (predictions location, data-lake root, verifier-calibration policy) changes. On the World of Logs corpus the agent runs under a \emph{family-held-out} leave-one-domain-out split: the dense bi-encoder is refit on the four training Apache projects (Spark, Cassandra, HBase, Flink) and the two held-out test families (Kafka, MariaDB-Server) are never seen during training. Both regimes are stricter generalisation tests than a same-distribution split.

\subsection{OpenTelemetry Demo: zero-shot to a structurally different application}
\label{sec:cross-app:otel}

\dsotel{} (\S\ref{sec:evaluation:datasets:otel}) is a polyglot synthetic application with fifteen-plus services, Kafka-mediated asynchronous communication, and a fifty-three-family scenario taxonomy more than twice the size of \dsob{}'s. The zero-shot setup holds every learned component fixed from the \dsob{}-trained configuration and varies only the upstream dataset.

Table~\ref{tab:headline-otel} reports the agent's headline metrics on the $247$-window test split, of which $119$ are evaluable for retrieval.

\begin{table}[ht]
  \centering
  \caption{Headline metrics on \dsotel{} ($n=247$, $n_{\text{eval}}=119$). $95\%$ percentile CIs from $1{,}000$-resample bootstrap at seed $42$.}
  \label{tab:headline-otel}
  \small
  \begin{tabular}{lrr}
    \toprule
    Metric & Point & $95\%$ CI \\
    \midrule
    Hit@1                     & $0.6471$ & $[0.5641, 0.7411]$ \\
    Hit@5                     & $0.7563$ & $[0.6726, 0.8333]$ \\
    MRR                       & $0.6880$ & $[0.6132, 0.7715]$ \\
    Triage accuracy           & $0.8421$ & $[0.7976, 0.8866]$ \\
    Pages per incident        & $1.000$  & --- \\
    Suppressions fired        & $1$      & --- \\
    Distinct plan identifiers & $4 / 6$  & --- \\
    \bottomrule
  \end{tabular}
\end{table}

The cross-dataset reading is the comparison against Table~\ref{tab:headline-ob}. Hit@5 of $0.7563$ on \dsotel{} is essentially identical to $0.7583$ on \dsob{}, with overlapping confidence intervals, despite the dataset being structurally different and every learned component held fixed. Hit@1 is $0.04$ lower in absolute terms (a difference within the per-dataset CI width). Triage accuracy is marginally higher ($0.8421$ versus $0.8343$), again within noise. The agent's design transfers across applications without any per-application retraining.

\paragraphhead{Cost.} On \dsotel{} the agent saves $65.3\%$ of wall time, language-model tokens, and dollar-equivalent cost against the always-on cascade counterfactual---higher than \dsob{}'s $64.1\%$. The mechanism is the same: the cheap-path-confident gate fires on more windows on \dsotel{} because the dataset's triage-numeric confidence distribution is more bimodal. The cost-savings claim survives the zero-shot transfer.

\paragraphhead{Pareto curve.} The threshold sweep over $\{0.50, \ldots, 0.95\}$ produces an even flatter curve on \dsotel{} than on \dsob{}: every threshold produces identical Hit@K \emph{and} identical cost, because the bimodal cheap-path confidence distribution and the smaller evaluable subset ($n=119$) leave essentially no windows in the threshold-sensitive intermediate range. The production-recommendation pattern from \S\ref{sec:results:pareto} (a lower threshold for free savings) generalises in spirit; the cross-dataset-specific recommendation is to calibrate the threshold per dataset rather than to assume \dsob{}'s default carries over.

\paragraphhead{Ablation.} The skill-ablation grid on \dsotel{} replicates \dsob{}'s pattern: removing the numeric-feature classifier collapses triage accuracy by $\Delta = -0.158$ (paired CI $[-0.207, -0.117]$, $p < 10^{-4}$), and no other single-primitive ablation moves Hit@K. The numeric-feature finding is therefore confirmed independently on a second dataset at the same significance level. The on-demand evidence-gathering loop's ablation cells, by contrast, produce exactly zero per-window deltas on every metric: on \dsotel{} the loop fires on $25$ of $247$ windows ($10.1\%$, versus \dsob{}'s $17.8\%$), and on those $25$ windows the rerank step does not change the rank-one ranking. The ReAct loop's non-significant result on \dsob{} sharpens to an exact zero on \dsotel{}.

\paragraphhead{Failure-mode catalog.} The tool-use catalog on \dsotel{} reports $25$ invocations of each tool. The pod-event tool returns an empty result on $22$ of $25$ ($88\%$) because the dataset's Kubernetes-event capture is sparser than \dsob{}'s; the other three tools succeed on all or all-but-two windows. The five-mode taxonomy (\S\ref{sec:results:failures}) thus produces a different per-tool distribution on \dsotel{}, but the same three never-firing modes (hallucinated, looping, budget-exhausted) remain at zero, confirming that the rule-based controller's structural guarantees hold across applications.

\subsection{Real Apache Jira: family-held-out leave-one-domain-out}
\label{sec:cross-app:wol}

The second transfer test is the paper's strongest external-validity claim. \dswol{} (\S\ref{sec:evaluation:datasets:wol}) replaces synthetic humanised tickets with $9{,}341$ real Apache Jira tickets from seven open-source projects, holds out two entire project families (Kafka and MariaDB-Server) from training, and evaluates on the held-out projects' tickets paired with telemetry-style query text. The dense bi-encoder is refit on the four training families (Spark, Cassandra, HBase, Flink) and tested on $1{,}648$ windows from the held-out families.

Table~\ref{tab:headline-wol} reports the agent's headline metrics.

\begin{table}[ht]
  \centering
  \caption{Headline metrics on \dswol{}, real Apache Jira tickets ($n = 1{,}648$, $n_{\text{eval}} = 48$ for agent retrieval, $276$ for cascade retrieval).}
  \label{tab:headline-wol}
  \small
  \begin{tabular}{lr}
    \toprule
    Metric & Point \\
    \midrule
    Hit@1                      & $\mathbf{0.958}$ \\
    Hit@5                      & $\mathbf{1.000}$ \\
    MRR                        & $\mathbf{0.976}$ \\
    Triage accuracy (default)  & $0.164$ \\
    Triage accuracy (\texttt{dense\_only}) & $0.921$ \\
    Novel recall / precision   & $0.000 / 0.000$ (no novel emissions) \\
    Pages per incident         & $\mathbf{1.000}$ \\
    Suppressions fired         & $510$ \\
    Distinct plan identifiers  & $2 / 6$ \\
    \bottomrule
  \end{tabular}
\end{table}

The retrieval headlines on \dswol{} are the highest of any dataset tested. Hit@1 of $0.958$, Hit@5 of $1.000$, and MRR of $0.976$ all exceed both synthetic benchmarks by margins outside the latter datasets' confidence intervals. The cross-project, real-production-language test is in fact \emph{easier} for the agent on the retrieval axis than the two synthetic benchmarks---in part because engineer-written ticket prose is denser in fault-specific terminology than humanised synthetic prose, and in part because the dataset's evaluable subset is a higher-precision sample under the time-ordered visibility rule.

\paragraphhead{The evaluable-subset asymmetry.} \dswol{} reports two evaluable counts. The cascade-evaluable subset is $276$: windows that carry at least one time-visible gold ticket. The agent-evaluable subset is $48$: windows on which the agent's cheap-path-confident gate did \emph{not} close (i.e.\ the agent escalated into the retriever stack at all). On the other $228$ windows the cheap-path triage was confident enough that the agent emitted a triage decision without invoking the retrievers, and so retrieval Hit@K is not defined there. The smaller agent-evaluable subset is therefore a consequence of the cost-saving gate from \S\ref{sec:results:cost}, not a sampling artefact, and the agent's $0.958$ Hit@1 is computed on the same windows where the cascade computes its $0.971$ on a wider gate-open subset.

\paragraphhead{Triage accuracy does not transfer at the default operating point.} The honest cross-dataset finding sits in the triage row of Table~\ref{tab:headline-wol}. At the agent's default configuration, triage accuracy on \dswol{} is $0.164$---substantially below both \dsob{} ($0.834$) and the test-split majority-class baseline ($0.496$). The mechanism is intelligible: \dswol{}'s test split has a $21$/$26$/$53$\% ticket-worthy / borderline / noise distribution (versus \dsob{}'s far higher ticket-worthy share), and the cheap-path triage stacker, trained on \dsob{}'s synthetic distribution, over-triages on \dswol{}'s text-only windows. A reduced \texttt{dense\_only} configuration that disables the over-confident escalation path recovers triage accuracy to $0.921$, which is well above the majority-class baseline and confirms that the architecture itself is sound. The cross-dataset claim the paper supports is therefore precise: retrieval generalises across the synthetic-to-real boundary without per-dataset triage refit; triage requires the operating point to be selected per-dataset.

\paragraphhead{Plan diversity is dataset-shaped.} The controller emits two distinct plan identifiers across the $1{,}648$ test windows, both within the active-fault branch. The reason is the dataset's structural shape: real Jira tickets do not carry the fault-window taxonomy (\texttt{pre\_fault\_baseline}, \texttt{recovery\_window}, \texttt{observation\_window}) that drives four of the six controller branches, so those branches cannot fire. The active-fault branch does fire, and the dataset's multi-ticket incident clusters (linked via Jira's \texttt{is-duplicate-of} and \texttt{Cloners} edges) do exercise the state-suppression branch: $510$ suppression events fire across the test split, the largest count of any of the three datasets, and the pages-per-incident ratio converges to $1.000$ over $901$ multi-ticket cases. The controller's branching machinery is exercised on \dswol{}; only the branches whose triggering conditions are dataset-specific (pre-fault, recovery, observation) cannot be.

\paragraphhead{ReAct on \dswol{}.} The dataset's text-only structure means three of the four on-demand tools (Kubernetes events, extended trace, Prometheus snapshot) auto-drop because their required capability flags are absent. The peer-incident tool is structurally applicable but never fires either, because the cheap-path gate closes on the BiEncoder's high-confidence retrieval and the controller does not open the tool-call gate. The failure-mode catalog records exactly zero tool invocations across the $1{,}648$ traces, and every tool-subset ablation produces an exact-zero paired delta on every metric. The cross-dataset ReAct picture is therefore consistent across all three datasets: a non-significant point-estimate lift on \dsob{}, exact zeros on \dsotel{} and \dswol{}. The framework is implemented; the empirical lift at the agent's current operating point is below the noise floor.

\paragraphhead{The verifier is structurally skipped on \dswol{}.} The dataset profile lists \dswol{} as \emph{known-harmful} for the language-model verifier, citing prior measurement that the verifier degrades Hit@5 on real Apache Jira tickets. The capability gate (\S\ref{sec:approach:gating}) refuses to invoke the verifier on \dswol{} as a structural property of the configuration, not a runtime decision; a unit-test assertion in the evaluation harness fires if the verifier ever appears in any \dswol{} trace's invocation list. This structural-defence pattern is what lets a measured negative result on one dataset (the verifier degrades retrieval on \dswol{}) become a configuration entry the controller honours on subsequent runs without re-measurement.

\subsection{What generalises across all three datasets}
\label{sec:cross-app:generalises}

Table~\ref{tab:generalises} summarises the cross-dataset pattern for each headline claim of the paper.

\begin{table*}[t]
  \centering
  \caption{Cross-dataset pattern. ``Generalises'' indicates the claim survives the dataset change with the same direction and similar magnitude; ``dataset-specific'' indicates a meaningful difference whose mechanism is identified in the corresponding section.}
  \label{tab:generalises}
  \small
  \begin{tabular}{llll}
    \toprule
    Claim                                                 & \dsob{}              & \dsotel{}             & \dswol{}              \\
    \midrule
    Cost savings vs.\ always-on cascade                   & $64.1\%$             & $65.3\%$              & verifier disabled by profile \\
    Pages per incident                                    & $1.000$              & $1.000$               & $1.000$               \\
    \texttt{triage\_numeric} essentiality ($p$)           & $<10^{-4}$           & $<10^{-4}$            & not applicable (no numeric flag) \\
    Plan diversity (distinct plan IDs)                    & $4$ of $6$           & $4$ of $6$            & $2$ of $6$ \\
    Suppression events fired                              & $20$                 & $1$                   & $510$                 \\
    Naive-baseline lift on Hit@5 (vs.\ BM25)              & $8.66\times$         & $1.34\times$          & $1.64\times$ ($245\times$ at Hit@1) \\
    \midrule
    \multicolumn{4}{l}{\emph{Dataset-specific:}} \\
    Headline Hit@1                                        & $0.6888$             & $0.6471$              & $\mathbf{0.958}$ \\
    Headline Hit@5                                        & $0.7583$             & $0.7563$              & $\mathbf{1.000}$ \\
    Triage accuracy (default config)                      & $0.8343$             & $0.8421$              & $0.164$ (recoverable to $0.921$) \\
    ReAct point-estimate lift on Hit@1                    & $+0.0151$ ($p=0.18$) & $0.000$               & $0.000$ \\
    Extended-trace tool harm                              & $\Delta=-0.018$, $p=0.002$ & flat (tool fires rarely)              & not applicable (no trace flag) \\
    False-novelty rate                                    & $40.6\%$             & $39.7\%$              & $\mathbf{0.0\%}$ \\
    \bottomrule
  \end{tabular}
\end{table*}

Three claims generalise without qualification across all three datasets and constitute the paper's strongest cross-dataset findings. First, the cost-savings claim holds at $64$ to $65$ percent against the always-on cascade counterfactual on both telemetry-rich datasets; on \dswol{} the verifier is structurally skipped, so the question does not arise. Second, the page-suppression mechanism converges to pages per incident $= 1.000$ on all three datasets, exercised by $1$ to $510$ suppression events depending on each dataset's multi-window incident density---most heavily on \dswol{}, where Jira's own duplicate-link metadata produces $901$ multi-ticket clusters. Third, the agent dominates a BM25-only naive baseline on Hit@5 by margins ranging from $1.34\times$ on \dsotel{} to $8.66\times$ on \dsob{}, and by $245\times$ on \dswol{} at the rank-one cut where BM25's keyword overlap is insufficient to disambiguate same-project tickets.

Two more claims generalise within their applicable scope. The numeric-feature classifier's essentiality for triage accuracy reaches $p < 10^{-4}$ on the two datasets that carry numeric features (\dsob{}, \dsotel{}); on \dswol{} the classifier is structurally absent. The controller's plan-diversity claim holds in the active-fault branch on all three datasets, and the state-suppression branch fires on \dswol{} where Jira's duplicate-link metadata makes the suppression rule meaningfully exercisable.

\subsection{What does not generalise, and why}
\label{sec:cross-app:doesnt}

Four findings from \S\ref{sec:results} are dataset-specific in informative ways.

\paragraphhead{Triage accuracy at the default operating point.} The same triage stacker that reaches $0.83$--$0.84$ on the telemetry-rich synthetic datasets reaches only $0.164$ on real Apache Jira at its default configuration, below the test-split majority-class baseline. The mechanism is the cheap-path triage classifier's tendency to over-page on text-only windows whose distribution differs structurally from the synthetic training data. A \texttt{dense\_only} reduced configuration recovers triage accuracy to $0.921$, confirming that the architecture is sound but the operating point is dataset-specific. Per-dataset operating-point selection is required for the deployable system; the architecture itself is portable.

\paragraphhead{The on-demand evidence-gathering lift.} The point-estimate Hit@1 lift from the ReAct loop on \dsob{} is $+0.0151$ ($p = 0.18$, non-significant). On \dsotel{} and \dswol{} the lift is exactly zero. On \dsotel{} the tools fire on too few windows and over too sparse a per-corpus pool ($147$ tickets versus \dsob{}'s $347$) for the rerank step's overlap-based scoring to change rank-one rankings; on \dswol{} the cheap-path gate closes on BiEncoder's high-confidence retrieval and the tool-call gate never opens, so the catalog records zero invocations across $1{,}648$ traces. The framework is implemented and exercised; the empirical lift at the agent's current operating point requires either a denser corpus, a more discriminative scoring head, or a gate that opens more aggressively---all future work.

\paragraphhead{The extended-trace tool's significant harm.} The $-0.018$ Hit@1 harm at $p = 0.002$ from the trace tool (\S\ref{sec:results:ablation}) is \dsob{}-specific. On \dsotel{} the same ablation cell produces exact-zero per-window delta because the trace tool fires on too few windows ($25$ total) for an isolated-trace cell to surface a swap. On \dswol{} the tool does not fire at all because the trace-summary capability is absent. The harm finding is therefore detectable only on the dataset where the tool is exercised most heavily, but its mechanism (cross-family token noise on call-path service names) is dataset-agnostic and the production recommendation (disable the tool by default; enable per scenario family) is portable.

\paragraphhead{The false-novelty rate.} The largest cross-dataset finding of the paper sits in the last row of Table~\ref{tab:generalises}. The false-novelty rate---the fraction of decisions in which the agent flags the window as novel despite a memory-side gold being present---is $40.6\%$ on \dsob{}, $39.7\%$ on \dsotel{}, and $0.0\%$ on \dswol{}. The two synthetic datasets agree to within one percentage point; on the real-Jira dataset the rate falls to zero. This is the cleanest cross-dataset discovery in the work.

\subsection{The false-novelty mechanism: a language-model-induced blind spot}
\label{sec:cross-app:falsenovel}

The mechanism is the structural skip of the language-model verifier on \dswol{} (\S\ref{sec:cross-app:wol}). The novelty composition layer (\S\ref{sec:approach:primitives}) OR-combines three independent novelty signals: the verifier's emitted novelty flag, a maximum-retrieval-confidence threshold, and a learned per-window classifier conditioned on coarse window metadata. The verifier's signal is the one that does not exist on \dswol{} because the verifier itself does not run. On \dsob{} and \dsotel{} the verifier runs and contributes its signal; on the windows where the verifier-induced novelty flag is over-eager (the verifier sees only the top-$k$ retrieved candidates and judges that none of them match the hypothesised root cause), the OR composition fires and the agent emits a false-novelty decision. On \dswol{} the verifier-induced signal is structurally absent, the OR composition fires only when one of the other two signals does, and the false-novelty rate falls to zero.

We test the mechanism directly. On \dsob{}, the windows that fire the false-novelty bucket overlap by $94\%$ with the windows on which the verifier emits its own novelty flag, and by under $10\%$ with windows where either of the other two signals would fire on their own. The verifier is the source of the false-novelty signal on the two synthetic datasets, not the OR-composition's failure to reject the verifier's input.

The implication is the paper's clearest design recommendation for future work. The verifier's novelty-flag emission is a tunable: it can be replaced by a more conservative criterion (e.g.\ requiring two of the three novelty signals to agree rather than one), or it can be disabled selectively on operating points where the verifier is not known-helpful for novelty. The latter is a configuration change; the former is a one-line edit to the novelty composer's OR-disjunction. Either change would, by the same mechanism that produces \dswol{}'s zero false-novelty rate, collapse \dsob{} and \dsotel{} false-novelty by a similar factor. The full evaluation of the change is left to a follow-up. The finding itself is what \S\ref{sec:cross-app}'s replication makes possible: a single-dataset paper would not have surfaced it.

\subsection{Summary of the cross-dataset evidence}
\label{sec:cross-app:summary}

The agent's design transfers across applications and across the synthetic-to-real boundary without per-dataset architectural changes. The strongest cross-dataset claims are the universal pages-per-incident ratio of $1.000$, the cost-savings rate at $64$ to $65$ percent on both telemetry-rich datasets, the three-dataset confirmation that BM25 alone is insufficient (with a $245\times$ rank-one lift on real Apache Jira), and the paired-delta significance of the numeric-feature classifier on triage accuracy on every dataset that carries numeric features. The retrieval signal on real Apache Jira (Hit@1 $= 0.958$, Hit@5 $= 1.000$) exceeds both synthetic benchmarks, in part because real engineer-written ticket prose carries more discriminative signal per token than humanised synthetic prose.

Two findings emerge that a single-dataset evaluation would have missed. First, triage accuracy at the default operating point does not transfer to real Apache Jira; a reduced configuration recovers it cleanly, but per-dataset operating-point selection is required. Second, the agent's largest blind spot on the synthetic datasets---a $\sim$$40\%$ false-novelty rate---collapses to zero on real Apache Jira, identifying the language-model verifier's overly-eager novelty flag as the mechanism and a one-line composition change as the candidate fix. Both findings are direct products of the three-dataset evaluation regime; neither would be visible from \dsob{} alone.
