% Section 10 — Conclusion
% Page budget: ~0.5 pages in acmart sigconf two-column (~370 words).
% Three paragraphs: what we built, what we measured, what comes next.
%
% Source-of-truth verification (no fake info):
%   - Cost savings 64.1% / 65.3%: PAPER-FINDINGS.md L24, L162.
%   - WoL v2 Hit@1=0.958 / Hit@5=1.000 (n_eval=48): PAPER-FINDINGS L175-176.
%   - Pages per incident=1.000 on all three: PAPER-FINDINGS L144, L160, L181.
%   - 1,648 WoL v2 test windows: PAPER-FINDINGS L182.
%   - 7 Apache projects, Kafka+MariaDB held out: DOCS/docs8/QnA.md §5,
%     data/derived/global/2026-06-15-wol-real-v2-global/README.md.
%   - ReAct null on all 3 datasets: PAPER-FINDINGS L262-283.
%   - Extended-trace tool harm p=0.002: PAPER-FINDINGS L325.
%   - False-novelty collapse 40.6% / 39.7% / 0.0%: PAPER-FINDINGS L399-409.
%   - WoL v2 triage 0.164 vs majority baseline 0.496: PAPER-FINDINGS L179.

\section{Conclusion}
\label{sec:conclusion}

On-call engineers spend most of their incident-response time not detecting anomalies but recognising which past Jira ticket describes the current failure. This paper presented a capability-adaptive incident-triage agent that turns the corporate Jira archive into a searchable institutional memory and surfaces the right ticket in seconds. For each monitoring window the agent inspects the available evidence---numeric telemetry, structured logs, distributed traces, Kubernetes events, or text alone---and assembles a per-window diagnostic plan that invokes an expensive language-model verifier only when the cheap retrieval path is uncertain. Every decision is replayable from a per-window execution trace.

We evaluated the agent across three datasets spanning the synthetic-to-real spectrum and contributed two new synthetic incident-triage benchmarks that for the first time pair the full telemetry stack---Loki logs, Tempo traces, Prometheus metrics, Kubernetes pod events---with humanised Jira tickets in a shared timeline schema. The agent matches the always-on cascade's retrieval quality at $64.1\%$ lower dollar-equivalent cost on Online Boutique and $65.3\%$ on the OpenTelemetry Demo, converges to one paged ticket per incident on every dataset, and reaches Hit@1 $= 0.958$ with Hit@5 $= 1.000$ on the retrieval-evaluable subset of $1{,}648$ real Apache Jira test windows---the highest accuracy across any dataset and, to our knowledge, the first agentic-retrieval result reported on a real Apache Jira corpus. The OpenTelemetry Demo transfer is zero-shot from the Online-Boutique-trained configuration; the World of Logs transfer is family-held-out, with Kafka and MariaDB-Server unseen during training.

Three directions follow naturally from what the cross-dataset evidence revealed. \emph{Per-family tool selection:} the active evidence-gathering loop's Hit@1 contribution is not statistically distinguishable from zero on any of the three datasets, and one of its four tools is significantly harmful on synthetic data; a controller that learns which tools to call on which scenario family from the per-window trace data the agent already emits is the natural next architectural step. \emph{Novelty-composition refinement:} the agent's $\sim$$40\%$ false-novelty rate on synthetic data collapses to zero on real Apache Jira when the language-model verifier's novelty flag is structurally absent, identifying the OR-disjunction's verifier input as the source and a one-line soft-AND change as the candidate fix. \emph{Per-dataset operating-point calibration:} the architecture transfers across the synthetic-to-real boundary unmodified, but the cheap-path triage threshold trained on synthetic-microservices class distributions should be treated as a per-dataset hyperparameter rather than a portable default. Each direction is a configuration entry in the agent's existing layered architecture; each can be validated by the paired-bootstrap measurement harness the paper already uses.
